---
title: "Effects of a GnRH agonist on giant pouched rat female sex behavior"
date: "`r Sys.Date()`"
author:
  - name: Beverly Lo
    email: bwl39@cornell.edu
    affiliation:
      - Cornell
  - name: Bhupinder Singh
    email: bs256@cornell.edu
    affiliation: 
      - Vetschool
  - name: Amit Hanadari-Levi
    email: XXX
    affiliation: Cornell
#    footnote:
#        - 1
  - name: Jackie Berman
    email: jeb525@cornell.edu
    affiliation: Cornell
  - name: Alexandra Chaves
    email: amc565@cornell.edu
    affiliation: Cornell
  - name: Aumena Choudhry
    email: dogwood25@outlook.com
    affiliation: Cornell
  - name: Angela R. Freeman
    email: afreeman@salisbury.edu
    corresponding_author: yes
    affiliation: 
      - Cornell
      - Salisbury
  - name: Alexander G. Ophir
    email: ophir@cornell.edu
    affiliation: Cornell 
address:
  - code: Cornell
    address: Cornell University, Department of Psychology, Ithaca, NY, 14853
  - code: Vetschool
    address: Cornell University, Department of Veterinary Medicine, Ithaca, NY, 14853
  - code: Salisbury
    address: Salisbury University, Department of Biological Sciences, Salisbury, MD 21853
footnote:
  - code: 1
    text: "Equal contribution"
  - code: 2
    text: "Current email address: \\href{mailto:cat@example.com}{cat@example.com}"

abstract: |
  Vaginal patency is an indicator of sexual maturity; if the vagina is closed, the female is unable to copulate. Males should be particularly attentive to cues of sexual availability and females should advertise accordingly. The African giant pouched rat is unusual because females can have fully fused vaginal openings (i.e. vaginal non-patency) despite being considered ‘adults’ by other metrics. Moreover, some females reversibly close their vaginal openings. Thus, vaginal patency in the pouched rat is a ‘flexible’ reproductive state. In order to better understand the endocrinology and social behavior of the African pouched rat, we used a long-acting GnRH agonist (deslorelin) to reduce circulating estradiol and to potentially alter vaginal patency. We hypothesized the change to non-patency is caused by a reduction in estradiol. We further hypothesized scent marking behavior would differ based on female patency and circulating estradiol. We monitored patency in 16 females; half received deslorelin implants, and the other half received control implants with only vehicle. We assessed scent marking behavior in response to male urine prior to implantation, one week after implants were inserted, and 6 weeks after implantation. After 6 months of physical examination, there were no changes in patency, but the deslorelin treated females gained significantly more weight than the controls. We observed no changes in scent marking behavior. In the future, we plan to conduct hormonal assays and imaging via MRI to observe potential additional physiological effects of deslorelin.

acknowledgements: |
  The authors thank the CARE staff at Cornell, in particular Wayne MacCabe and Lawrence Carlisle, for their efforts caring for the pouched rats used in this study. We also thank Drs. Ned Place, Soon Hon Cheong, Michael Sheehan,  and (XXXX) for feedback and assistance on the project. Alexandra Molina assisted in hormonal analyses. This study was funded by the Army Research Office (Grant XXXXX to AGO), DARPA (XXX to AGO) and a generous Army Educational Outreach Program Research Grant which supported BL and AC. Virbac Ltd. graciously provided vehicle implants for control animals.
  
keywords:
  - key
  - dictionary
  - word
fontsize: 10pt
#spacing: halfline # could also be oneline
#classoptions:
#  - endnotes
bibliography:  ./library.bib
#csl: apa.csl
#link-citations: yes
output: rticles::oup_article
#urlcolor: orange
#linkcolor: green
#citecolor: red
#header-includes:
#  - \usepackage[nomarkers,tablesfirst]{endfloat} # For figures and tables at end
#  - \usepackage{lineno} # For line numbering
#  - \linenumbers # For line numbering
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("/Users/afreeman/Desktop")
getwd()

#load all libraries
library(ade4)
library(ade4TkGUI)
library(ggplot2)
library(ggthemes)
library(lme4)
library(lmerTest)
library(emmeans)
library(car)
library(cowplot)
library(dbplyr)
library(dplyr)
library(broom)
library("ggpubr")



## Function for standard error
serr <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))

##Theme for plots
theme_alex <- function(base_size=24, base_family ="Arial") {
  theme_few() %+replace%
    theme(axis.ticks = element_blank(), legend.title=element_text(size=24, face="bold"), 
          axis.title=element_text(size=24,face="bold"),
          axis.line = element_line(colour = "black"),
          panel.border = element_blank(),
          legend.position = "right")
  
}

COLOR1 <- "#D98DE9"
COLOR2 <- "#E9D98D"
```

# Introduction
Many animal species possess more highly developed chemical detection systems relative
to humans. One such species, and the focus of this research, is the African giant pouched rat (C. ansorgei). Due to their excellent sense of smell, relatively long life spans, and low maintenance costs, African giant pouched rats have become an important tool in leading demining efforts lead by APOPO (Anti-Persoonsmijnen Ontimijende Product Ontwikkeling; a non-governmental organization). However, despite their crucial role in humanitarian efforts, little is known about this species. Beyond knowledge about their excellent sense of smell, we have yet to mount significant research into understanding their development, social behaviors, and endocrinology. 

Understandably, if a female’s vagina is closed (i.e. nonpatent), the female is unable to mate. Typically, vaginal patency develops at the start of adulthood (around 21 d in mice and 40 d in rats) (Miller et al., 2002 and Mandl and Zuckerman, 1952). However, some rodents, including female African giant pouched rats, have a unique reproductive system in which they can exhibit delayed vaginal patency development well past adulthood (Weir, 1970 and Freeman et al., 2019). Females in species such as the cuis (Galea Musteloides) do not develop patency until they are within contact
of a male (Weir, 1970). Thus, delayed patency may correlate with the female’s sexual receptivity or be influenced by environmental factors, such as availability of potential mates (Whitsett et al., 1982). In pouched rats, vaginal patency can develop significantly later than in other rodent species, in some cases for several years (3 years or more) (Freeman et al., 2019). In some species with delayed patency, pouched rats included, females exhibit the ability to become reversibly nonpatent after reaching full adult maturity (Freeman et al., 2019). This has been linked to several possible mechanisms. Species in the Hysricomorpha suborder exhibit patency changes based on the estrus cycle (Touma, 2001). In both deer mice (Peromyscus maniculatus) and Siberian hamsters  (Phodopus sungorous), patency is dependent on photoperiod, indicating patency may be regulated by seasons (Whitsett et al., 1982; Place et al., 2004). In particular, Siberian hamsters have been shown to develop patency in response to estradiol induced uterine maturation as well, with nonpatent females yielding smaller uterine horns than patent females (Place et al., 2004). In pouched rats, magnetic resonance imaging (MRI) has yielded a similar finding: nonpatent females have smaller uterine horns compared to that of patent females (Freeman et al., in prep). However, pouched rats exhibit patency changes without exposure to males, photoperiod changes, or changes in the estrus cycle.

Vaginal patency, and female sexual receptivity more generally, can have a profound effect on various social behaviors, including male preferences (Freeman et al., 2019). Previous research on pouched rats is very limited, but efforts to understand their mate preferences have recently been reported. Specifically, patent females displayed a preference for odor derived from males with longer anogenital distances (AGD), whereas non-patent females showed no preference (Freeman et al., 2019). Longer AGD in male rodents is thought to correlate with general competitiveness (Delbarco-Trillo et al., 2009). Longer AGDs are related to exposure to androgens in utero, where increased levels of androgens lead to longer AGD (Even, 1992; Nonneman, 1992; Ryan, 2002; vom Saal, 1978). In pouched rats, males with longer AGD showed a distinct preference for vaginally patent females over nonpatent females, while males with shorter AGD showed no preference (Freeman et al., 2019). However, although we know that patency can impact preferences when females are receiving social olfactory signals, we do not know whether patency can influence female olfactory signaling behaviors.

Deslorelin, a long-acting gonadotropin releasing hormone (GnRH) agonist, is used in many species as a form of contraception to induce a pseudo-menopausal state (see Grosset, 2012; Munson, 2001). A GnRH agonist binds receptors located in the anterior pituitary and initially induces an increase in gonadotropin release, such as luteinizing hormone (LH) and folliclestimulating hormone (FSH). LH and FSH secretion stimulate the gonads to release sex steroid hormones (progesterone and estrogen for females). However, with prolonged use, the GnRH agonist will desensitize the anterior pituitary and induce negative feedback that eventually prevents the ovaries from releasing sex hormones (Grosset, 2012). In ferrets, deslorelin was subcutaneously
implanted and was considered a suitable method of contraception (Zeeland et al., 2014). Previous studies in female cats (queens) have shown that deslorelin implantation significantly reduced the levels of estradiol in the blood even without knowledge of the cat’s estrus stage (Munson, 2001). We chose a GnRH analogue for our experiment because it offered us the ability to measure endogenous circulating estradiol and progesterone while also fulfilling the need to have a long lasting and removable method to secrete GnRH. We have previously tried to measure serum
estradiol levels in pouched rats but were unable to collect accurate readings because the levels were too low (Allen et al., personal comm.). By using a GnRH agonist, we hoped to increase estradiol levels prior to long-term suppression via negative feedback in order to detect the circulating levels through hormone assays.

This study aimed to better understand the endocrinology and social behaviors of the
African pouched rat. We tested the hypothesis that reversion to reversible non-patency is mediated by a decrease in estradiol and that scent-marking behavior is influenced by the reproductive state of the female. We predicted that injection of deslorelin would lower estradiol levels in the long term and cause the vaginal membrane to fuse together, inducing nonpatency. We also predicted reversibly nonpatent females would have a lower frequency of scent marking behavior compared to when they were patent.



## Methods

### Animals
 The animals used in this study included animals captured at our field site in Morogoro, Tanzania as adults, and their descendants that were born in our captive colony.

We used 16 females for Deslorelin (or saline) implantation and scent marking behavior testing. Nine females were wild-caught of unknown age and the other 7 females were born in captivity as descendants from wild-caught rats. Captive-born animals were weaned from their mothers after 2 months and were housed individually, like the rest of the colony. We conducted baseline scent marking experiments (prior to Deslorlelin implantation) on 13 of these patent females. All the animals were housed singly in
standard rabbit cages (61 x 61 x 41 cm; Hoeltge Inc., Cincinnati, OH, U.S.A.), with males and females in separate rooms. Newspaper was used as bedding in each cage. The colony followed a 12:12 h light:dark cycle (lights off at 1400 hours Eastern Daylight Time, EDT) and was maintained at 21 °C and 45% humidity. The animals were fed Teklad F6 rodent diet (Envigo, Huntingdon, U.K.), provided water ad libitum, and given weekly treats of fruits, vegetables, or seeds. All of the work detailed here was approved by the U.S. Army Medical Research and Materiel Command (USAMRMC) Animal Care and Use Review Office (ACURO) and the Cornell Institutional Animal Care and Use Committee (IACUC).
 
### Implantation
16 adult females were briefly anesthetized using isoflurane gas and then implanted with either a Deslorelin implant (n=8, Virbac Ltd, Westlake, TX) or Vehicle implant (n=8, Virbac Ltd, Westlake, TX) placed subcutaneously on the dorsal side of the animal just caudal to the shoulder blades and near the midline. All females were patent at the start of the experiment, except one that had reverted to non-patency one week prior to implantation. Animals were placed into two groups of 8 (4 Deslorelin and 4 Control in each group). All implants were place on the Monday of 'week 1' (Figure 1) for each group.

### Blood Collection

We collected blood for serum for (future) hormone analysis. Blood was collected on
Monday, Wednesday, and Friday on weeks 1, 2, and 6 between 9am and 12pm (Figure 1). Due to time constraints, we split the females into two groups, each with 4 control and 4 experimental females, and designated 6 of the total 12 weeks for blood collection. Approximately 1 mL of blood was collected through the tail vein using a 22g needle and syringe and was allowed to clot at room temperature. After at least one hour, we spun the blood at 2100 rpm for 30 minutes at room temperature and pipetted off the serum into a microcentrifuge tube. As needed, blood was respun at 2100 rpm for 5 minutes and any additional serum was added to the sample before storage in a -80°C freezer. 

### Physical examination

We closely monitored vaginal patency of the female African giant pouched rats on a weekly basis with visual inspection assisted by a periodontal probe when the female was under anesthesia. We scored the measurements as either nonpatent (i.e., fused, with no opening in the skin), patent (i.e., unfused, with opening in the skin), or patent closed (i.e., unfused, with partial opening in the skin). While females were anesthetized, we weighed the rats and recorded nipple size and color. Nipple size was measured using a wire gauge measuring tool that included numerical measurements with decreasing diameter indicated by larger numbers. Nipple color was compared to the body color and recorded as bright pink, medium pink, or pale pink (Figure 2).

### Behavior
Scentarking trials were completed before implantation while females were patent, and at week 2 and 6 (1 and 5 weeks post implantation, respectively). These dates were chosen to coincide with predicted changes in circulating estradiol.

Male choice tests were conducted near 6 months post implantation. 



### Results

#Patency
We detected no effect of GnRH agonist on patency after 6 months of treatment. 

# Nipples
We detected no effect of Deslorelin on nipple size

Deslorelin treatment did result in mass gain, as predicted.

# Scent marking
There was no effect of Deslorelin on the total investigation of the male mark, or on the number of scent marks or their size.

# Behavior in social choice test

Deslorelin treatment significantly impacted female behavior and male behavior towards the treated female.

```{r, echo=FALSE}

beverly <- read.csv("Beverlydata.csv")

AMITdata <-read.csv("BEVdata.csv")
#summary(AMITdata)
AMITdata$Trial <- as.factor(AMITdata$Trial)
AMITdata$Sex <- as.factor(AMITdata$Sex)
AMITdata$ID <- as.factor(AMITdata$ID)
AMITdata$Status <- as.factor(AMITdata$Status)
AMITdata$PercentAGG <- AMITdata$Aggression.dur/AMITdata$Total.time.w.male*100
AMITdata$PercentPOS <- AMITdata$Sniff.dur/AMITdata$Total.time.w.male*100

#Remove one trial where both females were saline-treated (human error)
AMITdata <-AMITdata[!(AMITdata$Trial=="1510959"),] 
```

```{r, echo=FALSE, warning = FALSE}
#head(beverly) #this will show you the first few lines to make sure it's loaded
#summary(beverly) #make sure there are all three timepoints, 13 pre, 16 week 2, 16 week 6

#The following are the linear models and anovas
beverly$ID <- as.factor(beverly$ID)

#sniffing duration
model.sniff <- lmer(Sniff_dur ~ Treatment*Timepoint +(1|ID), data=beverly)
summary(model.sniff)
anova(model.sniff)

#licking the scent
model.lick <- lmer(Lickscent_dur ~ Treatment*Timepoint +(1|ID), data=beverly)
summary(model.lick)
anova(model.lick)

#saliva/chewing area
model.salarea <- lmer(Saliva_area ~ Treatment*Timepoint +(1|ID), data=beverly)
summary(model.salarea)
anova(model.salarea)

#urine area
model.urarea <- lmer(Urine_area ~ Treatment*Timepoint +(1|ID), data=beverly)
summary(model.urarea)
anova(model.urarea)

#Anogenital rubbing
model.agrub <- lmer(AG_rubbing_dur~ Treatment*Timepoint +(1|ID), data=beverly)
summary(model.agrub)
anova(model.agrub)

#grooming duration
model.groom <- lmer(Groom_dur ~ Treatment*Timepoint +(1|ID), data=beverly)
summary(model.groom)
anova(model.groom)

#for count data, we use negative binomial distributions, 
#since these are not 'normal' distribution data
model.urcount <- glmer.nb(Urine_count ~ Treatment*Timepoint +(1|ID), data=beverly)
summary(model.urcount)
anova(model.urcount)

#this is for post-hoc testing

emmip(model.urcount,Treatment~Timepoint)
emmeans(model.urcount, pairwise~Treatment|Timepoint)

#for count data, we use negative binomial distributions, 
#since these are not 'normal' distribution data
model.salcount <- glmer.nb(Saliva_count ~ Treatment*Timepoint +(1|ID), data=beverly)
summary(model.salcount)
anova(model.salcount)

emmip(model.salcount,Treatment~Timepoint)
emmeans(model.salcount, pairwise~Treatment|Timepoint)

## make an investigation variable
beverly$investigate <- beverly$Lickscent_dur + beverly$Sniff_dur

#anova for investigation
model.invest <- lmer(investigate ~ Treatment*Timepoint +(1|ID), data=beverly)
summary(model.invest)
anova(model.invest)


## Following for boxplots
ggboxplot(beverly, x = "Timepoint", y = "investigate", 
          fill = "Treatment", palette = c("#ADD8E6", "#00688B"),
          lwd=1,
          ylab = "Duration of investigation (s)", xlab = "Timepoint")+
  theme_alex()


ggboxplot(beverly, x = "Timepoint", y = "Urine_area", 
          fill = "Treatment", palette = c("#ADD8E6", "#00688B"),
          lwd=1,
          ylab = "Area of urine mark (cm2)", xlab = "Timepoint")+
  theme_alex()


ggboxplot(beverly, x = "Timepoint", y = "Groom_dur", 
          fill = "Treatment", palette = c("#ADD8E6", "#00688B"),
          lwd=1,
          ylab = "Grooming duration (s)", xlab = "Timepoint")+
  theme_alex()


ggboxplot(beverly, x = "Timepoint", y = "Urine_count", 
          fill = "Treatment", palette = c("#ADD8E6", "#00688B"),
          lwd=1,
          ylab = "Urine marking counts", xlab = "Timepoint")+
  theme_alex()

ggboxplot(beverly, x = "Timepoint", y = "AG_rubbing_dur", 
          fill = "Treatment", palette = c("#ADD8E6", "#00688B"),
          lwd=1,
          ylab = "Anogenital rubbing duration (s)", xlab = "Timepoint")+
  theme_alex()

#following for line plots
ggplot(data = beverly, aes(x = Timepoint, y = investigate, group = ID, color = Treatment, fill=Treatment)) + 
  geom_line() + 
  geom_point(shape = 21, size = 3, color="black") + 
  xlab("Test Time") + 
  ylab("Investigation time (s)") + 
  scale_fill_manual(values = c("#ADD8E6", "#00688B")) +
  scale_color_manual(values = c("#ADD8E6", "#00688B")) +
  theme_alex()
ggsave("Investigation (sniff and lick).png", width=8, height=6, dpi=600)


ggplot(data = beverly, aes(x = Timepoint, y = Groom_dur, group = ID, color = Treatment, fill=Treatment)) + 
  geom_line() + 
  geom_point(shape = 21, size = 3, color="black") + 
  xlab("Test Time") + 
  ylab("Grooming duration (s)") + 
  scale_fill_manual(values = c("#ADD8E6", "#00688B")) +
  scale_color_manual(values = c("#ADD8E6", "#00688B")) +
  theme_alex()
ggsave("Self grooming.png", width=8, height=6, dpi=600)


ggplot(data = beverly, aes(x = Timepoint, y = Urine_area, group = ID, color = Treatment, fill=Treatment)) + 
  geom_line() + 
  geom_point(shape = 21, size = 3, color="black") + 
  xlab("Test Time") + 
  ylab("Urine area (cm2)") + 
  scale_fill_manual(values = c("#ADD8E6", "#00688B")) +
  scale_color_manual(values = c("#ADD8E6", "#00688B")) +
  theme_alex()
ggsave("Urine area.png", width=8, height=6, dpi=600)



ggplot(data = beverly, aes(x = Timepoint, y = Urine_count, group = ID, color = Treatment, fill=Treatment)) + 
  geom_line() + 
  geom_point(shape = 21, size = 3, color="black") + 
  xlab("Test Time") + 
  ylab("Urine mark counts") + 
  scale_fill_manual(values = c("#ADD8E6", "#00688B")) +
  scale_color_manual(values = c("#ADD8E6", "#00688B")) +
  theme_alex()
ggsave("Urine mark counts.png", width=8, height=6, dpi=600)

```



```{r, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
### Statistics
#1. Examine whether treatment affected interaction.
Males <- AMITdata %>% filter(Sex=="M") #prevent double counting datapoints.

#Because each male was only in one trial, lmer and lm are equivalent.
fit1 <- lmer(Total.time.w.male ~ Status + (1|Trial), data=Males)
anova(fit1)
fit1a <- summary(fit1)
lm(Total.time.w.male ~ Status, data=Males)
InteractingAnova <- anova(lm(Total.time.w.male ~ Status, data=Males))

#2. Does status predict aggression time?
fit2 <- lmer(PercentAGG ~ Status * Sex + (1|Trial), data=AMITdata)
anova(fit2)
fit2.emm <- emmeans(fit2,  ~  Status*Sex)
#qqnorm(resid(fit2))
#qqline(resid(fit2))

#3. Does status predict sniffing time?
fit3 <- lmer(PercentPOS ~ Status * Sex + (1|Trial), data=AMITdata)
anova(fit3)
fit3.emm <- emmeans(fit3,  ~  Status*Sex)
#qqnorm(resid(fit3))
#qqline(resid(fit3))
```
*The time the male and Saline-treated female spent interacting together was greater than the time the male and Deslorelin-treated female spent together.*
Mean time spent interacting Deslorelin and male: Control and male: , t = `r round(fit1a$coefficients[2,4], digits= 3)`, p = `r round(fit1a$coefficients[2,5], digits= 4)`.
 however, this was no longer significant after corrections for multiple comparisons (adjusted $\alpha$ = 0.001)



```{r, echo=FALSE}



### First plot, Male aggression towards females ###
plotdata <- AMITdata %>% filter(ID=="Male") %>% group_by(Status) %>% mutate(meanagg = mean(Aggression.dur), stanaggerr = serr(Aggression.dur))  

aggplot1 <- ggplot(plotdata, aes(x=Status, y=Aggression.dur, fill= Status)) +
  scale_y_continuous(expand= c(0,0), limits = c(0, 40)) +
  #coord_cartesian(ylim = c(0, 10)) +
  geom_jitter(aes(color = (as.factor(Status))), size = 2, position=position_dodge(width=0.9), alpha=0.5) +
  geom_bar(aes(color = as.factor(Status)), alpha = .5,
           width=0.8,
           stat="summary", size = 1, position="dodge") +
  geom_errorbar(aes(ymin=meanagg-stanaggerr, ymax=meanagg+stanaggerr, color = as.factor(Status)), width=.4,
                position=position_dodge(.9)) +
  labs(x = "Implant Status", y="Male aggression towards females \n duration (s) ± SE")+ #Axes labels
  theme_alex() +
  scale_fill_manual(values=c(COLOR1, COLOR2)) + #colors in hexadecimal 
  scale_color_manual(values=c("Black", "Black")) + #colors in hexadecimal 
  # annotate("segment", x = 1, xend = 4, y = 40, yend = 40, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2.5, y = 41.5, size =6, label = "**") +
  #annotate("segment", x = 1, xend = 3, y = 35, yend = 35, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2, y = 36.5, size =6, label = "*") +
  guides(fill="none", color = "none") +
  theme(legend.title=element_text(size=8, face="bold"), 
        legend.text = element_text(size=8),
        axis.text = element_text(size=7, colour="black"),
        axis.title=element_text(size=8,face="bold"),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        legend.position = "right"
  )
aggplot1 #show plot

##### How much did males spend sniffing females ###

plotdata2 <- AMITdata %>% filter(ID=="Male") %>% group_by(Status) %>% mutate(meansniff = mean(Sniff.dur), stanSnifferr = serr(Sniff.dur))  

sniffplot1 <- ggplot(plotdata2, aes(x=Status, y=Sniff.dur, fill= Status)) +
  scale_y_continuous(expand= c(0,0), limits = c(0, 250)) +
  #coord_cartesian(ylim = c(0, 10)) +
  geom_jitter(aes(color = (as.factor(Status))), size = 2, position=position_dodge(width=0.9), alpha=0.5) +
  geom_bar(aes(color = as.factor(Status)), alpha = .5,
           width=0.8,
           stat="summary", size = 1, position="dodge") +
  geom_errorbar(aes(ymin=meansniff-stanSnifferr, ymax=meansniff+stanSnifferr, color = as.factor(Status)), width=.4,
                position=position_dodge(.9)) +
  labs(x = "Implant Status", y="Male sniffing towards females \n duration (s) ± SE")+ #Axes labels
  theme_alex() +
  scale_fill_manual(values=c(COLOR1, COLOR2)) + #colors in hexadecimal 
  scale_color_manual(values=c("Black", "Black")) + #colors in hexadecimal 
  # annotate("segment", x = 1, xend = 4, y = 40, yend = 40, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2.5, y = 41.5, size =6, label = "**") +
  #annotate("segment", x = 1, xend = 3, y = 35, yend = 35, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2, y = 36.5, size =6, label = "*") +
  guides(fill="none", color = "none") +
  theme(legend.title=element_text(size=8, face="bold"), 
        legend.text = element_text(size=8),
        axis.text = element_text(size=7, colour="black"),
        axis.title=element_text(size=8,face="bold"),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        legend.position = "right"
  )
sniffplot1 #show plot

### Association time ###

plotdataINTERACT <- AMITdata %>% filter(ID=="Male") %>% group_by(Status) %>% mutate(meaninteract = mean(Total.time.w.male), stanInteracterr = serr(Total.time.w.male))  

INTERACTplot <- ggplot(plotdataINTERACT, aes(x=Status, y=Total.time.w.male, fill= Status)) +
  scale_y_continuous(expand= c(0,0), limits = c(0, 1000)) +
  #coord_cartesian(ylim = c(0, 10)) +
  geom_jitter(aes(color = (as.factor(Status))), size = 2, position=position_dodge(width=0.9), alpha=0.5) +
  geom_bar(aes(color = as.factor(Status)), alpha = .5,
           width=0.8,
           stat="summary", size = 1, position="dodge") +
  geom_errorbar(aes(ymin=meaninteract-stanInteracterr, ymax=meaninteract+stanInteracterr, color = as.factor(Status)), width=.4,
                position=position_dodge(.9)) +
  labs(x = "Implant Status", y="Time spent interacting (s) ? SE")+ #Axes labels
  theme_alex() +
  scale_fill_manual(values=c(COLOR1, COLOR2)) + #colors in hexadecimal 
  scale_color_manual(values=c("Black", "Black")) + #colors in hexadecimal 
  # annotate("segment", x = 1, xend = 4, y = 40, yend = 40, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2.5, y = 41.5, size =6, label = "**") +
  #annotate("segment", x = 1, xend = 3, y = 35, yend = 35, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2, y = 36.5, size =6, label = "*") +
  guides(fill="none", color = "none") +
  theme(legend.title=element_text(size=8, face="bold"), 
        legend.text = element_text(size=8),
        axis.text = element_text(size=9, colour="black"),
        axis.title=element_text(size=9,face="bold"),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        legend.position = "right"
  )
INTERACTplot #show plot
ggsave("Interaction Plot.png", width=4, height=4, dpi=600)

### Female Aggression ####
plotdataF1 <- AMITdata %>% filter(Sex=="F") %>% group_by(Status) %>% mutate(meanagg = mean(Aggression.dur), stanaggerr = serr(Aggression.dur))  

aggplot2 <- ggplot(plotdataF1, aes(x=Status, y=Aggression.dur, fill= Status)) +
  scale_y_continuous(expand= c(0,0), limits = c(0, 40)) +
  #coord_cartesian(ylim = c(0, 10)) +
  geom_jitter(aes(color = (as.factor(Status))), size = 2, position=position_dodge(width=0.9), alpha=0.5) +
  geom_bar(aes(color = as.factor(Status)), alpha = .5,
           width=0.8,
           stat="summary", size = 1, position="dodge") +
  geom_errorbar(aes(ymin=meanagg-stanaggerr, ymax=meanagg+stanaggerr, color = as.factor(Status)), width=.4,
                position=position_dodge(.9)) +
  labs(x = "Implant Status", y="Female aggression towards male \n duration (s) ± SE")+ #Axes labels
  theme_alex() +
  scale_fill_manual(values=c(COLOR1, COLOR2)) + #colors in hexadecimal 
  scale_color_manual(values=c("Black", "Black")) + #colors in hexadecimal 
  # annotate("segment", x = 1, xend = 4, y = 40, yend = 40, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2.5, y = 41.5, size =6, label = "**") +
  #annotate("segment", x = 1, xend = 3, y = 35, yend = 35, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2, y = 36.5, size =6, label = "*") +
  guides(fill="none", color = "none") +
  theme(legend.title=element_text(size=8, face="bold"), 
        legend.text = element_text(size=8),
        axis.text = element_text(size=7, colour="black"),
        axis.title=element_text(size=8,face="bold"),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        legend.position = "right"
  )
aggplot2 #show plot

##Female Sniffing###

plotdataF2 <- AMITdata %>% filter(Sex=="F") %>% group_by(Status) %>% mutate(meansniff = mean(Sniff.dur), stanSnifferr = serr(Sniff.dur))  

sniffplotF1 <- ggplot(plotdataF2, aes(x=Status, y=Sniff.dur, fill= Status)) +
  scale_y_continuous(expand= c(0,0), limits = c(0, 250)) +
  #coord_cartesian(ylim = c(0, 10)) +
  geom_jitter(aes(color = (as.factor(Status))), size = 2, position=position_dodge(width=0.9), alpha=0.5) +
  geom_bar(aes(color = as.factor(Status)), alpha = .5,
           width=0.8,
           stat="summary", size = 1, position="dodge") +
  geom_errorbar(aes(ymin=meansniff-stanSnifferr, ymax=meansniff+stanSnifferr, color = as.factor(Status)), width=.4,
                position=position_dodge(.9)) +
  labs(x = "Implant Status", y="Female sniffing towards male \n duration (s) ± SE")+ #Axes labels
  theme_alex() +
  scale_fill_manual(values=c(COLOR1, COLOR2)) + #colors in hexadecimal 
  scale_color_manual(values=c("Black", "Black")) + #colors in hexadecimal 
  # annotate("segment", x = 1, xend = 4, y = 40, yend = 40, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2.5, y = 41.5, size =6, label = "**") +
  #annotate("segment", x = 1, xend = 3, y = 35, yend = 35, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2, y = 36.5, size =6, label = "*") +
  guides(fill="none", color = "none") +
  theme(legend.title=element_text(size=8, face="bold"), 
        legend.text = element_text(size=8),
        axis.text = element_text(size=7, colour="black"),
        axis.title=element_text(size=8,face="bold"),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        legend.position = "right"
  )
sniffplotF1 #show plot

##
plot_grid(INTERACTplot, NULL, aggplot1, aggplot2, nrow=2, rel_widths = c(2, 2))
ggsave("Aggression Panel.png", width=8, height=3, dpi=600) #save panel locally

plot_grid(INTERACTplot, NULL, sniffplot1, sniffplotF1, nrow=2, rel_widths = c(2, 2))
ggsave("Sniff Panel.png", width=8, height=3, dpi=600) #save panel locally


### Percentage plots ##

plotPositive <- AMITdata %>% filter(Sex=="F") %>% group_by(Status) %>% mutate(mean1 = mean(PercentPOS), stanerr1 = serr(PercentPOS))  

plotPOS <- ggplot(plotPositive, aes(x=Status, y=PercentPOS, fill= Status)) +
  scale_y_continuous(expand= c(0,0), limits = c(0, 100)) +
  #coord_cartesian(ylim = c(0, 10)) +
  geom_jitter(aes(color = (as.factor(Status))), size = 2, position=position_dodge(width=0.9), alpha=0.5) +
  geom_bar(aes(color = as.factor(Status)), alpha = .5,
           width=0.8,
           stat="summary", size = 1, position="dodge") +
  geom_errorbar(aes(ymin=mean1-stanerr1, ymax=mean1+stanerr1, color = as.factor(Status)), width=.4,
                position=position_dodge(.9)) +
  labs(x = "Implant Status", y="Percent of Interaction time \n with female sniffing")+ #Axes labels
  theme_alex() +
  scale_fill_manual(values=c(COLOR1, COLOR2)) + #colors in hexadecimal 
  scale_color_manual(values=c("Black", "Black")) + #colors in hexadecimal 
  # annotate("segment", x = 1, xend = 4, y = 40, yend = 40, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2.5, y = 41.5, size =6, label = "**") +
  #annotate("segment", x = 1, xend = 3, y = 35, yend = 35, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2, y = 36.5, size =6, label = "*") +
  guides(fill="none", color = "none") +
  theme(legend.title=element_text(size=8, face="bold"), 
        legend.text = element_text(size=8),
        axis.text = element_text(size=7, colour="black"),
        axis.title=element_text(size=8,face="bold"),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        legend.position = "right"
  )
plotPOS #show plot



plotAggressive <- AMITdata %>% filter(Sex=="F") %>% group_by(Status) %>% mutate(mean1 = mean(PercentAGG), stanerr1 = serr(PercentAGG))  

plotAGG <- ggplot(plotAggressive, aes(x=Status, y=PercentAGG, fill= Status)) +
  scale_y_continuous(expand= c(0,0), limits = c(0, 30)) +
  #coord_cartesian(ylim = c(0, 10)) +
  geom_jitter(aes(color = (as.factor(Status))), size = 2, position=position_dodge(width=0.9), alpha=0.5) +
  geom_bar(aes(color = as.factor(Status)), alpha = .5,
           width=0.8,
           stat="summary", size = 1, position="dodge") +
  geom_errorbar(aes(ymin=mean1-stanerr1, ymax=mean1+stanerr1, color = as.factor(Status)), width=.4,
                position=position_dodge(.9)) +
  labs(x = "Implant Status", y="Percent of Interaction time \n with female aggression")+ #Axes labels
  theme_alex() +
  scale_fill_manual(values=c(COLOR1, COLOR2)) + #colors in hexadecimal 
  scale_color_manual(values=c("Black", "Black")) + #colors in hexadecimal 
  # annotate("segment", x = 1, xend = 4, y = 40, yend = 40, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2.5, y = 41.5, size =6, label = "**") +
  #annotate("segment", x = 1, xend = 3, y = 35, yend = 35, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2, y = 36.5, size =6, label = "*") +
  guides(fill="none", color = "none") +
  theme(legend.title=element_text(size=8, face="bold"), 
        legend.text = element_text(size=8),
        axis.text = element_text(size=9, colour="black"),
        axis.title=element_text(size=9,face="bold"),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        legend.position = "right"
  )
plotAGG #show plot
ggsave("Aggression Plot2.png", width=4, height=4, dpi=600) #save panel locally

### Percentage plots male ##


plotPositiveM <- AMITdata %>% filter(Sex=="M") %>% group_by(Status) %>% mutate(mean1 = mean(PercentPOS), stanerr1 = serr(PercentPOS))  

plotPOSM <- ggplot(plotPositiveM, aes(x=Status, y=PercentPOS, fill= Status)) +
  scale_y_continuous(expand= c(0,0), limits = c(0, 100)) +
  #coord_cartesian(ylim = c(0, 10)) +
  geom_jitter(aes(color = (as.factor(Status))), size = 2, position=position_dodge(width=0.9), alpha=0.5) +
  geom_bar(aes(color = as.factor(Status)), alpha = .5,
           width=0.8,
           stat="summary", size = 1, position="dodge") +
  geom_errorbar(aes(ymin=mean1-stanerr1, ymax=mean1+stanerr1, color = as.factor(Status)), width=.4,
                position=position_dodge(.9)) +
  labs(x = "Implant Status", y="Percent of Interaction time \n with male sniffing")+ #Axes labels
  theme_alex() +
  scale_fill_manual(values=c(COLOR1, COLOR2)) + #colors in hexadecimal 
  scale_color_manual(values=c("Black", "Black")) + #colors in hexadecimal 
  # annotate("segment", x = 1, xend = 4, y = 40, yend = 40, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2.5, y = 41.5, size =6, label = "**") +
  #annotate("segment", x = 1, xend = 3, y = 35, yend = 35, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2, y = 36.5, size =6, label = "*") +
  guides(fill="none", color = "none") +
  theme(legend.title=element_text(size=8, face="bold"), 
        legend.text = element_text(size=8),
        axis.text = element_text(size=7, colour="black"),
        axis.title=element_text(size=8,face="bold"),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        legend.position = "right"
  )
plotPOSM #show plot



plotAggressiveM <- AMITdata %>% filter(Sex=="M") %>% group_by(Status) %>% mutate(mean1 = mean(PercentAGG), stanerr1 = serr(PercentAGG))  

plotAGGM <- ggplot(plotAggressiveM, aes(x=Status, y=PercentAGG, fill= Status)) +
  scale_y_continuous(expand= c(0,0), limits = c(0, 30)) +
  #coord_cartesian(ylim = c(0, 10)) +
  geom_jitter(aes(color = (as.factor(Status))), size = 2, position=position_dodge(width=0.9), alpha=0.5) +
  geom_bar(aes(color = as.factor(Status)), alpha = .5,
           width=0.8,
           stat="summary", size = 1, position="dodge") +
  geom_errorbar(aes(ymin=mean1-stanerr1, ymax=mean1+stanerr1, color = as.factor(Status)), width=.4,
                position=position_dodge(.9)) +
  labs(x = "Implant Status", y="Percent of Interaction time \n with male aggression")+ #Axes labels
  theme_alex() +
  scale_fill_manual(values=c(COLOR1, COLOR2)) + #colors in hexadecimal 
  scale_color_manual(values=c("Black", "Black")) + #colors in hexadecimal 
  # annotate("segment", x = 1, xend = 4, y = 40, yend = 40, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2.5, y = 41.5, size =6, label = "**") +
  #annotate("segment", x = 1, xend = 3, y = 35, yend = 35, colour = "black", size=1, alpha=1) +
  #annotate("text", x = 2, y = 36.5, size =6, label = "*") +
  guides(fill="none", color = "none") +
  theme(legend.title=element_text(size=8, face="bold"), 
        legend.text = element_text(size=8),
        axis.text = element_text(size=9, colour="black"),
        axis.title=element_text(size=9,face="bold"),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        legend.position = "right"
  )
plotAGGM #show plot
ggsave("Aggression Panel percent2.png", width=4, height=4, dpi=600)

plot_grid(INTERACTplot, NULL, plotAGG, plotAGGM, nrow=2, rel_widths = c(2, 2))
ggsave("Aggression Panel percent.png", width=8, height=3, dpi=600) #save panel locally

plot_grid(INTERACTplot, NULL, plotPOS, plotPOSM, nrow=2, rel_widths = c(2, 2))
ggsave("Sniff Panel percent.png", width=8, height=3, dpi=600) #save panel locally



```



## Discussion

# References
